---
title: "DataVisualisationAssignment2"
author: "aungmyatphyo22221184"
date: "2023-03-14"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(sf)
library(dplyr)
library(lubridate)
library(scales)
library(ggplot2)
library(tidyr)
library(colorspace)
library(colorblindr)

setwd("~/Downloads")
file1<-"/Users/aungphyo/Downloads/DV2/CovidCountyStatisticsIreland_v2.shp"
IRL_counties_Covid19 <- st_read(file1, quiet = TRUE)

#filter the data of all counties on 2021,12,21
#normalize DailyCCase and Confirmed C columns per 100,000 population
q1<- IRL_counties_Covid19%>%
  filter(TimeStamp == ymd("2021-12-21"))%>%
  mutate(ConfirmedC_per_100k =  round(100000  * ConfirmedC/Population,1))%>%
  mutate(DailyCCase_per_100k =  round(100000  * DailyCCase/Population, 1))

#colour variable
cb<-desaturate(palette_OkabeIto,0.25)
cb1<-cb[c(1,2)]
#new variable for difference colour between galway and others
Counties=c('Other','Other','Other','Other','Other',
           'Other','Galway','Other','Other','Other',
           'Other','Other','Other','Other','Other','Other',
           'Other','Other','Other','Other','Other','Other',
           'Other','Other','Other','Other')
#bind a new variable Counties column to q1 dataframe
q1<-cbind(q1,Counties)

q11<-ggplot(q1,
            #order by ascending value of confirmedC per 100k value
            aes(x=reorder(CountyName,ConfirmedC_per_100k),
                y=ConfirmedC_per_100k,color=Counties))+
  geom_point(size=3.5)+
  ggtitle('Cumulative Cases per 100k population per county on 21st Dec 2021 ')

q11<-q11+theme(panel.spacing = unit(1,'lines'))

#change the y label name and discard x label
q11<-q11+xlab('')+
  ylab('Cumulative no of cases per 100,000 population')

#change the scales of y axis and add comma after 3 digits
q11<-q11+scale_y_continuous(breaks = seq(10000,19000,by=1000),
                            limits = c(10000,19000),
                            labels = comma)

q11<-q11+scale_fill_manual(values=cb1)

#theme
q11<-q11+theme_bw()+theme(
  panel.background = element_rect(fill = "white"),
  panel.grid.major = element_line(size = 0.1, 
                                  linetype = 'solid',
                                  colour = "lightgrey"), 
  #remove x axis major grid line
  panel.grid.major.x = element_blank(),
  panel.grid.minor = element_line(size = 0.075, 
                                  linetype = 'solid',
                                  colour = "lightgrey"),
  #reduce the text size of x and angle into 30
  axis.text.x = element_text(angle=30,vjust=1,hjust=1,size=6),
  #bold the title name
  plot.title = element_text(face='bold',size=9),
  axis.title.y = element_text(size=7),
  axis.line.y = element_blank(), 
  axis.line.x = element_blank(), 
  panel.border = element_blank()
 )


q11
```

For question 1, my hand drawing graph & ggplot graph are similar. Firstly, there is a difference in y-axis scale upper limits (19000) to cover highest data points and text style from 10k to 10,000. In ggplot graph, there are major and minor y axis grid lines which means easier to read the data. For colour, colour blind package is adapted for considering about the people with colour problem. For legend, Counties is described as legend title and 2 categories such as Galway and other which will help the reader for better understanding this graph. Finally, the plot title is modified.






```{r}
library(sf)
library(dplyr)
library(lubridate)
library(scales)
library(ggplot2)
library(tidyr)
library(colorspace)
library(colorblindr)

setwd("~/Downloads")
file1<-"/Users/aungphyo/Downloads/DV2/CovidCountyStatisticsIreland_v2.shp"
IRL_counties_Covid19 <- st_read(file1, quiet = TRUE)

#filter the data of all counties on 2021,12,21
#normalize DailyCCase and Confirmed C columns per 100,000 population
q2<- IRL_counties_Covid19%>%
  filter(TimeStamp == ymd("2021-12-21"))%>%
  mutate(ConfirmedC_per_100k =  round(100000  * ConfirmedC/Population,1))%>%
  mutate(DailyCCase_per_100k =  round(100000  * DailyCCase/Population, 1))

#colour variable
cb<-desaturate(palette_OkabeIto,0.25)
cb1<-cb[c(1,2)]

#mean value of ConfirmedC per 100k column
m<-mean(q2$ConfirmedC_per_100k)
#mean 13528.98

#to differ the colours among below the mean counties,
#above the mean counties and galway (3 colours)
mm<-c('a','a','b','b','a','a','g','b','a',
      'b','b','b','a','a','a','b','a','a',
      'b','b','b','b','a','a','b','b')
#bind new variable mm column to dataframe q2
q2<-cbind(q2,mm)


q22<-ggplot(q2,
            #order by ascending values of ConfirmedC per 100k value
            aes(x=reorder(CountyName,ConfirmedC_per_100k),
               y=ConfirmedC_per_100k,colour=mm))+
  #draw a mean horizontal line by colour dark blue
  geom_hline(yintercept = m, 
             linetype="solid",
             size=0.2,
             colour='#00008B')+
  #to show divergence from mean horizontal line
  geom_segment(aes(x=reorder(CountyName,ConfirmedC_per_100k),
                   y=m,
                   xend=CountyName, 
                   yend=ConfirmedC_per_100k),
               size=0.3,
               show.legend = FALSE)+
  geom_point(size=2)+
  #title name
  ggtitle("Divergence from mean cumulative cases per 100k on 21st,Dec,2021")+
  #caption explanation
  labs(caption = "horizontal dark blue line represents mean number of cases (13528.98)")
  

#adding manual color and labels
q22<-q22+scale_colour_manual(values= c("#32cd32","#ce5a6c","212f85"), 
                             name = "",
                      labels = c("Above_the_mean","Below_the_mean","Galway")) 
#theme
q22<-q22+theme(panel.spacing = unit(1,'lines'))+
  #remove x label
  xlab('')+
  #change y label name
  ylab('Cumulative no of cases per 100,000 population')+
  #change the scales of y axis and add comma after 3 digits
  scale_y_continuous(breaks = seq(10000,19000,
                                  by=1000),
                     limits = c(10000,19000),
                     labels = comma)+
  theme_bw()+theme(
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(size = 0.1, 
                                    linetype = 'solid',
                                    colour = "lightgrey"),
    #remove x axis major grid line
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_line(size = 0.075, 
                                    linetype = 'solid',
                                    colour = "lightgrey"),
    #reduce the text size of x and angle into 30
    axis.text.x = element_text(angle=30,
                               vjust=1,
                               hjust=1,
                               size=6),
    #bold the title name
    plot.title = element_text(face='bold',
                              size=9),
    axis.title.y = element_text(size=7),
    axis.line.y = element_blank(), 
    axis.line.x = element_blank(), 
    panel.border = element_blank()
  )
  
q22
```

For question2, both graphs are similar. Firstly, there is a difference in y-axis scale upper limits (19000) to cover highest data points and text style from 10k to 10,000. In ggplot graph, there are major and minor y axis grid lines which means easier to read the data. The solid dark blue mean line is drawn rather than dotted ones and green colour for above the mean counties, red for below the mean counties and blue for galway is applied. Each dot means individual values and vertical sticks in the lollipop plot describes how much to differ from the mean value. Moreover, a legend is added and there is a caption about mean line below the graph. These features will provide better visualisation. Finally, the plot title is modified.





```{r}
library(sf)
library(dplyr)
library(lubridate)
library('ggplot2')
library('tidyr')
library(scales)

setwd("~/Downloads")
file1<-"/Users/aungphyo/Downloads/DV2/CovidCountyStatisticsIreland_v2.shp"
IRL_counties_Covid19 <- st_read(file1, quiet = TRUE)

#filter the County Wicklow data
county<-subset(IRL_counties_Covid19,CountyName=="Wicklow")
#filter the 18th week duration between 1st Jan 2021 to 6th May 2021
Wicklow<- county%>%
  filter(TimeStamp >=as.Date('2021-01-01')& TimeStamp<=as.Date('2021-05-06'))


#plot with x axis timestamp and y axis DailyCCase
q3<-ggplot(data =Wicklow, aes(x =TimeStamp, y = DailyCCase)) +
  #daily data during 18th week
  geom_point(alpha = 0.2, colour='red',
             size = 1,
             na.rm = TRUE) +
  #title
  ggtitle('Daily Number of Confirmed Cases in County Wicklow over a 18-week period')+
  #explain 18th week duration
  labs(caption = "18th weeks are from 1st,Jan,2021 to 6th,May,2021")

#change y label and discard X label
q3<-q3+labs(x = " ", 
            y = "Daily Covid Case")

#change x label into week and breaks as 1 week
q3<-q3+scale_x_date(
  name = 'Week',
  breaks = "1 week",
  labels = date_format("%W"),
  expand = c(0.01, 0))

#to show trends, draw the blue line
q3<-q3+geom_line(color = "blue", 
                 size = .5) 

q3<-q3+theme(panel.spacing = unit(1,'lines'))

#change the y scales
q3<-q3+scale_y_continuous(breaks = seq(0,140,
                                       by=10),
                          limits = c(0,142))

#theme
q3<-q3+theme_bw()+theme(
  panel.background = element_rect(fill = "white"),
  panel.grid.major = element_line(size = 0.1, 
                                  linetype = 'solid',
                                  colour = "lightgrey"), 
  panel.grid.minor = element_line(size = 0.075, 
                                  linetype = 'solid',
                                  colour = "lightgrey"),
  #remove minor x and y grid lines
  panel.grid.minor.y = element_blank(),
  panel.grid.minor.x = element_blank(),
  #bold the title
  plot.title = element_text(face='bold',
                            size=9),
  axis.title.x = element_text(size=7),
  axis.title.y = element_text(size=7),
  axis.line.y = element_blank(), 
  axis.line.x = element_blank(), 
  panel.border = element_blank())

q3  
```


For question 3, both graphs are similar. Firstly, y-axis upper limit scales is changed into 140 and x-axis scales are described as 1,2,3 and x-axis label is as Week. In ggplot graph, smoothing process is not applied because the data is not too big or too complexed for smoothing. Major gridlines of x-axis and y-axis are applied for better visualisation. For colour, line is drawn as blue and dots on each data is as red. In these difference colour, blue line shows the trend and each red dots serve as the individual data value. Finally, the plot title is modified and a caption is added to mention about the 18th week period.